---
title: "TxDb: Genes, Transcripts, and Genomic Locations"
author: 
- name: Sean Davis
  affiliation: Center for Cancer Research, National Cancer Institute, NIH
  email: seandavi@gmail.com
package: GenomicDataIntegration
output: 
  BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r init, include=FALSE}
library(knitr)
opts_chunk$set(message=FALSE, cache=TRUE)
```

```{r style, echo = FALSE, results = 'asis'}
BiocStyle::markdown()
```


The `r Biocpkg("GenomicFeatures")` package retrieves and manages transcript-related features from
the UCSC Genome Bioinformatics[^ucsc]
and BioMart2 data resources. The package is
useful for ChIP-chip, ChIP-seq, and RNA-seq analyses.

[^ucsc]: https://genome.ucsc.edu

```{r library}
library('GenomicFeatures')
```
                               
# Annotation

```{r gencodegtf}
download.file("https://www.encodeproject.org/files/gencode.v24.primary_assembly.annotation/@@download/gencode.v24.primary_assembly.annotation.gtf.gz", "gencode.v24.primary_assembly.annotation.gtf.gz")
```

```{r txdb}
library(GenomicFeatures)
txdb = makeTxDbFromGFF('gencode.v24.primary_assembly.annotation.gtf.gz')
```

To save the `txdb` database for later and avoid having to recreate it 
every time we use it, we can use `saveDb()` and, later, `loadDb()`

```{r savetxdb}
library(AnnotationDbi)
saveDb(txdb, 'txdb.gencode24.sqlite')
```

```{r exploreTxDb}
txdb = loadDb(file = 'txdb.gencode24.sqlite')
genes(txdb)
```

# Expression Data




